Computing fusion products of MV cycles using the Mirkovi\'c--Vybornov isomorphism

### 2021-06-13 submission.tex
changed some delimiters

%!TEX program = lualatex
% 
\documentclass{article} 
\usepackage{basic}
\usepackage{float}
\title{Computing fusion products of MV cycles using the Mirkovi\'c--Vybornov isomorphism}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
\date{\today}

@@

\maketitle
% \tableofcontents 
\begin{abstract}
    The fusion of two Mirkovi\'c--Vilonen cycles is a degeneration of their product, defined using the Beilinson--Drinfeld Grassmannian.  In this paper, we put in place a conceptually elementary approach to computing this product in type $A$.  We do so by transferring the problem to a fusion of generalized orbital varieties using the Mirkovi\'c--Vybornov isomorphism.  As an application, we explicitly compute all cluster exchange relations in the coordinate ring of the upper-triangular subgroup of $ \GL_4$, confirming that all the cluster variables are contained in the Mirkovi\'c--Vilonen basis.
\end{abstract}

\section{Introduction}
\label{s:intro}
% 
\subsection{Geometric Satake and Mirkovi\'c--Vilonen cycles}
\label{ss:gsmv}
% 
The geometric Satake equivalence of Mirkovi\'c and Vilonen \cite{mirkovic2007geometric} is an equivalence of categories between the category of spherical perverse sheaves on the affine Grassmannian of a reductive group $ G $ and the category of representations of its Langlands dual group $G^\vee$.  This foundational result provides a powerful tool to study representation theory using geometry.  In particular, under this equivalence, the MV cycles (short for Mirkovi\'c--Vilonen) in the affine Grassmannian of $ G $ index bases for irreducible representations of $ G^\vee $. In this paper we work with $ G = \GL_m$ and identify $ G^\vee = \GL_m$ as well.

In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov supply a geometric version of symmetric and skew Howe $(\GL_m,\GL_N)$ dualities.  They relate Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $N\times N$ nilpotent orbits on the one hand, and to Nakajima quiver varieties on the other hand.    The second author \cite{dthesis} showed  that under the Mirkovi\'c--Vybornov isomorphism, the MV cycles are identified with certain varieties of matrices, called generalized orbital varieties.

Essential to the geometric Satake equivalence, the BD Grassmannian (short for Beilinson--Drinfeld) is used to define a fusion product on the category of spherical perverse sheaves on the affine Grassmannian.  In \cite{anderson2003polytope}, Anderson used the BD Grassmannian to define a fusion of MV cycles.  He constructed a family where the general fibre is a Cartesian product of two MV cycles, and the special fibre, called the fusion, is a (non-reduced) union of MV cycles.  Anderson conjectured that this fusion product matches the multiplication in the coordinate ring of the unipotent subgroup $ N \subset\GL_m$.  This conjecture was proven in \cite{baumann2019mirkovic}, where it was also established that the MV cycles give a basis for $ \CC[N]$.

\subsection{Fusion using generalized orbital varieties}
\label{ss:fugov}
% 
Computing the fusion product of MV cycles using the BD Grassmannian is quite difficult, both computationally and conceptually.  A few such computations were done by Anderson and Kogan in \cite{anderson2006algebra} in a somewhat ad-hoc fashion.  Another perspective on these computations (using convolution Grassmannians) was recently studied by Baumann, Gaussent and Littelmann \cite{baumann2020bases}.

@@

\begin{theorem}
    Let $ Z', Z'' $ be MV cycles.  The structure constants for the of the corresponding MV basis vectors in $\CC[N]$ are equal to the intersection multiplicities of generalized orbital varieties in the fusion of the generalized orbital varieties corresponding to $ Z', Z''$.
\end{theorem}

@@

\subsection{Cluster algebras and MV cycles}
One motivation for computing products of MV basis elements is to relate the cluster algebra structure on $\CC[N]$ to the MV basis. Cluster algebras were introduced by Fomin and Zelevinsky in \cite{fomin2002cluster1} and it was shown in \cite{geiss2007initial}, using a construction of \cite{berenstein2005cluster3}, that $\CC[N]$ has a natural cluster algebra structure.

This structure starts with an initial seed $\Sigma = (\{x_1,\dots,x_r\},B)$, where $x_i \in \CC[N]$ are a cluster of variables, $B$ is a certain matrix, and $ r = m(m-1)/2$.
We obtain new variables through a process called mutation, where for a mutable variable $x_i$, its mutation is the unique element $x_i^* \in \CC[N]$ such that 
$$
    x_ix_i^* = x_+ + x_- \,.
$$
This equation is called an exchange relation, and the terms $x_+$ and $x_-$ are certain monomials in the $x_j$, $j\neq i$,  determined by $B$. 
It gives us the new cluster $\{x_1, \dots, x_i^*, \dots, x_r\}$.

The initial cluster $\{x_1,\dots,x_r\}$ as well as all possible variables that are obtained through successive mutations are known as cluster variables. 
The cluster monomials are products of cluster variables that are supported on a single cluster.

There are two well-known bases of $\CC[N]$ whose relationship with the cluster structure has been heavily studied: the dual canonical basis of \cite{lusztig1990canonicalbases} and the dual semicanonical basis of \cite{lusztig2000semicanonical}.
In \cite{kang2018monoidal}, Kang, Kashiwara, Kim and Oh showed that the cluster monomials are contained in the dual canonical basis, while, in \cite{geiss2006rigid}, Gei\ss, Leclerc and Schr\"oer proved that they are all contained in the dual semicanonical basis.
% \cite[Theorem~2.8(3)]{geiss2006rigid}.
%

The MV basis, the dual canonical basis, and the dual semicanonical basis are all examples of biperfect bases of $\CC[N]$ \cite[Section 2]{baumann2019mirkovic}. In particular, they share a crystal structure.
In \cite[Appendix]{baumann2019mirkovic}, we showed that the MV basis differs from the dual semicanonical basis at the same point in the underlying crystal as where the dual semicanonical basis differs from the dual canonical basis.  
% 
The fact that this point at which each of these bases differ is not a cluster monomial in any of these bases, gives evidence for the following conjecture of Anderson and Kogan \cite[Conjecture 5.1]{anderson2006algebra}.
\begin{conjecture}\label{conj:cluster in MV}
    The cluster monomials are contained in the MV basis. 
\end{conjecture}

The conjecture is true in type $A_n$ for $n\leq 3$ since in these cases $\CC[N]$ has a unique biperfect basis \cite[Section 2.3]{baumann2019mirkovic}.
% 
More generally, cluster variables called flag minors, coming from initial seeds constructed in \cite[Definitions~2.2--2.3]{berenstein2005cluster3}, belong to each biperfect basis of $\CC[N]$ \cite[Remark~2.10]{baumann2019mirkovic}. 
And recently in \cite[Proposition 7.2]{baumann2020bases} cluster monomials supported on clusters of flag minors satisfying a certain combinatorial condition where shown to be in the MV basis.

Our hope is that each exchange relation can be realized using the fusion of MV cycles. This would imply that all cluster variables are in the MV basis. In \Cref{ss:GL4 examples} we verify that this is indeed the case in type $A_3$ by directly computing each fusion product corresponding to an exchange relation.

\section{Supporting Cast}
\label{s:players}
% 
\subsection{Rings and discs}
\label{ss:rings}
Set $\cO = \cO_0 = \CC\xt$ and $\cK = \cK_0 = \CC\xT$.
We will also consider $\cO_s = \CC\xt[t-s]$ and its fraction field $\cK_s=\CC\xT[t-s]$, for any $ s \in \CC\setminus\{0\} $, as well as $ \Oinf = \CC\xt[t^{-1}] $ and $\Kinf = \CC\xT[t^{-1}]$. For any point $ s \in \PP = \PP^1$, $ \cO_s$ is the completion of the local ring $ \mathcal O_{\PP, s} $ and thus the formal spectrum of $ \cO_s$ is the formal neighbourhood $ D_s$ of $ s $ (also called the formal disc centered at $ s$).  Similarly, the formal spectrum of the field $\cK_s$ is the deleted formal neighbourhood (or punctured disc), denoted $ D_s^\times$.

Note that for $ s \in \AA = \AA^1$, we have an obvious isomorphism  $\cO_s \cong \cO$ and $\cK_s\cong\cK$ taking $ t-s $ to $ t$. 
%

\subsection{Groups}
\label{ss:groups}
% 
Let $H $ be an algebraic group over $ \CC $.  We will be interested in $ H(R)$ where $ R $ is a $\CC$-algebra, for example $R = \CC[t], \cO_s$, etc.  
% 
Note that evaluation at $ t = s$ provides a group homomorphism $ H(\cO_s) \rightarrow H$.  We denote the kernel of this map by $ H_1(\cO_s)$, often called the  first congruence subgroup.  We will be particularly interested in this construction in the case $ s = \infty$, which gives us the group $ H_1(\Oinf)$.

Throughout the paper, we fix $ m \in \NN$.  We let $G = \GL_m$ and we let $T\subset G$ be the maximal torus of diagonal matrices. Let $ N, N_- \subset G$ denote the subgroups of upper and lower triangular matrices with 1s on the diagonal. 
We identify $\ZZ^m$ with the coweight lattice of $G$ and say that a coweight $\nu = (\nu_1,\dots,\nu_m)$ is \new{dominant} if $\nu_1\ge\cdots\ge\nu_m$ and \new{effective} if $\nu_j\ge 0$ for all $j$.
% 
If $\nu$ is both effective and dominant, then it is a partition of size $|\nu| = \nu_1 + \cdots + \nu_m\in\NN$.  Let $ Q_+ \subset \ZZ^m$ denote the positive root cone, explicitly we have
$$
    Q_+ = \{ (\nu_1, \dots, \nu_m) : \nu_1 + \cdots + \nu_j \ge 0 \text{ for $j = 1, \dots, m-1 $ and } \nu_1 + \dots + \nu_m = 0 \}\,. 
$$
We define a partial order on $ \ZZ^m$ by $ \lambda \ge \mu $ if and only if $ \lambda - \mu \in Q_+ $.

Given $s\in\CC$ we define $(t-s)^\nu$ to be the diagonal matrix 
\[
\begin{bmatrix}
    (t-s)^{\nu_1} \\
    & (t-s)^{\nu_2} \\ 
    & & \ddots \\
    & & & (t-s)^{\nu_m}
\end{bmatrix} 
\]
which we can view in $G(K)$ for any ring $ K $ containing $(t-s)^{-1}$ and $\CC[t]$. For example, $(t-s)^\nu\in G(K)$ for $ K = \Ks$ or $\CC(t)$.

We will also be interested in the affine space $ M_m$ of $m\times m$ matrices.  Note that for any $ \CC$-algebra $ R $, $G(R)$ consists of those matrices $ g \in M_m(R) $ whose determinant is invertible in $ R$. Thus for example $ (t-s)^\mu\in M_m(\CC[t])$ for all effective $ \mu \in \ZZ^m$ but $(t-s)^\mu\in G(\CC[t])$ if and only if $ \mu = 0 $.
%

\subsection{Lattices}
\label{ss:lat}
% 
We will use the lattice model for the affine Grassmannian, so it is useful to recall the following definition. Let $ R \subset K$ be two $\CC$-algebras (usually, but not always, $K$ will be a field). 
Consider $ K^m $ as a $K$-module. By restriction $ K^m$ can be viewed as an $R$-module.  An \new{$R$-lattice} in $K^m$ is an $R$-submodule $ L \subset K^m$ which is a free $R$-module of rank $ m $ and satisfies $ L \otimes_R K = K^m $. Equivalently, $ L = \Sp_R(v_1, \dots, v_m)$ where $v_1, \dots, v_m$ are free generators of $K^m$. 
%

$ R^m \subset K^m $ is called the standard lattice. The group $\GL_m(K) $ acts transitively on the set of $R$-lattices in $K^m$, thus giving a bijection between this set and $\GL_m(K)/\GL_m(R)$, since $ \GL_m(R) $ is the stabilizer of the standard lattice.

We will be particularly interested in $\CC[t]$-lattices in $ \CC(t)^m$. Given such a lattice $ L $ and a point $ a \in \CC$, the \new{specialization} of $ L $ at $ a $ is the lattice in $\cK_a^m$ defined as $ L(a) := L \otimes_{\CC[t]} \cO_a $. 
If $L(a) = \cO_a^m$ then $L$ is said to be \new{trivial at $a$}. 
For example, the lattice $(t-s)^{-1}\CC[t] \subset \CC(t)$ is trivial at any $a\ne s$, since $ t-s $ is invertible in $ \cO_a$.

\section{Affine Grassmannians} 
\label{s:affgrs}
% 
We now define various versions of the affine Grassmannian which will play important roles in this paper. 
% 
Each definition is made group-theoretically and then restated as a moduli space of vector bundles and as a moduli space of lattices.
% 
We also sketch how to pass between descriptions.

In these definitions, $ \Vtriv$ denotes the trivial rank $ m $ vector bundle.

\begin{definition}
\label{def:gr}
     The \new{ordinary affine Grassmannian} $\Gr = G(\cK)/G(\cO)$.
\end{definition}    
It is the moduli space of vector bundles with trivializations,
$$
\Gr = 
    \left\{ 
        (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_0 $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D_0^\times $} 
    \right\} \,. 
$$
It is also a moduli space of lattices, 
$$ 
\Gr = 
    \left\{ L \subset \cK^m : \text{ $L$ is a $\cO$-lattice} \right\}\,.
$$
We obtain a lattice from a pair $ (V,\varphi) $ by setting $ L = \Gamma(D_0, V)$ which is embedded into $ \cK^m = \Gamma(D_0^\times, \Vtriv)$ using $ \varphi$.  
On the other hand, to get a lattice from the group-theoretic description $ G(\cK) / G(\cO) $ we set $ L = g \cO^m$ for $ g \in G(\cK)$.
% 
\begin{definition}
\label{def:grs}
    For any $ s \in \CC $, the \new{ordinary affine Grassmannian} $\Gr_s = G(\cK_s)/G(\cO_s)$ \new{at} $ s $. 
\end{definition}  
% 
As above, we have modular and lattice descriptions:
% 
\begin{gather*}
\Gr_s = 
    \left\{ (
        V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_s $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D^\times_s $} 
    \right\}\,, \\
\Gr_s = 
    \left\{ 
        L \subset \cK_s^m : \text{ $L$ is a $\cO_s$-lattice} 
    \right\} \,. 
\end{gather*}
% 
\begin{definition}
\label{def:grth}
    The \new{thick affine Grassmannian} $\Grth = G(\Kinf)/G(\CC[t])$.
\end{definition}
Again, we have the following modular and lattice descriptions:
\begin{gather*}
\Grth = 
    \left\{ 
        (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ \PP $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D_\infty $ } 
    \right\}\,, \\
\Grth = 
    \left\{ 
        L : L \subset  \Kinf^m \text{ is a $\CC[t]$-lattice} 
    \right\}\,.
\end{gather*}
% 
\begin{definition} 
\label{def:bdgr}
    The two point \new{Beilinson--Drinfeld Grassmannian} $$\pi : \Grbd\to \AA$$ with one point fixed at 0, and the second point $s\in\AA$ varying.
\end{definition}
% 
It is described in modular terms by 
$$
\Grbd = 
    \left\{ 
        (V,\varphi,s) : V\text{ is a rank $m$ vector bundle on }\PP, \varphi : V \xrightarrow{\sim} \Vtriv \text{ on } \PP \setminus \{0, s\}  
    \right\}\,. 
$$
% 
The fibre of $\Grbd \to \AA$ over $ s \in \AA $ will be denoted $ \Gr_{0,s} $ and is given by
% 
$$ 
    G(\CC[t, t^{-1}, (t-s)^{-1}]/G(\CC[t])\,. % := \pi^{-1}(s) 
$$
We also have the lattice descriptions:
% 
\begin{gather*}
\Grbd = 
    \left\{ 
        (L,s) : L \subset  \CC(t)^m\text{ is a $\CC[t]$-lattice trivial at any }a \ne 0, s 
    \right\}\,, \\
% , $L$ is 
\Gr_{0,s} = 
    \left\{ 
        L : L \subset  \CC[t,t^{-1},(t-s)^{-1}]^m \text{ is a $\CC[t]$-lattice} 
    \right\}\,.
\end{gather*}
% 
\begin{definition} 
\label{def:grplus}
The \new{positive part} of $\Gr $, resp.\ $\Grth$, is defined by 
\begin{gather*}
    \Gr^+ = \left(M_m(\cO) \cap G(\cK)\right) / G(\cO)\,,\,\, 
    \text{resp.\ } {\Grth}^+ = \left(M_m(\CC[t]) \cap G(\Kinf)\right) / G(\CC[t])\,. 
\end{gather*}
\end{definition}
% 
In modular terms, $\Gr^+$ (resp.\ ${\Grth}^+$) is the set of those $ (V, \varphi)$ where $ \varphi : V \rightarrow \Vtriv $ extends to an inclusion of coherent sheaves over $ D_0 $ (resp.\ over $ \PP$).

In lattice terms, $ \Gr^+$ (resp.\ ${\Grth}^+$) contains those lattices $L$ which are contained in $\cO^m$ (resp.\ $\CC[t]^m$).
%

>>>>

% 
\documentclass{article} 
\usepackage{basic}
\usepackage{float}
\title{Computing fusion products of MV cycles using the Mirkovi\'c--Vybornov isomorphism}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
\date{\today}

@@

\maketitle
% \tableofcontents 
\begin{abstract}
    The fusion of two Mirkovi\'c--Vilonen cycles is a degeneration of their product, defined using the Beilinson--Drinfeld Grassmannian.  In this paper, we put in place a conceptually elementary approach to computing this product in type \(A\).  We do so by transferring the problem to a fusion of generalized orbital varieties using the Mirkovi\'c--Vybornov isomorphism.  As an application, we explicitly compute all cluster exchange relations in the coordinate ring of the upper-triangular subgroup of \( \GL_4\), confirming that all the cluster variables are contained in the Mirkovi\'c--Vilonen basis.
\end{abstract}

\section{Introduction}\label{s:intro}
% 
\subsection{Geometric Satake and Mirkovi\'c--Vilonen cycles}\label{ss:gsmv}
% 
The geometric Satake equivalence of Mirkovi\'c and Vilonen~\cite{mirkovic2007geometric} is an equivalence of categories between the category of spherical perverse sheaves on the affine Grassmannian of a reductive group \( G \) and the category of representations of its Langlands dual group \(G^\vee \).  This foundational result provides a powerful tool to study representation theory using geometry.  In particular, under this equivalence, the MV cycles (short for Mirkovi\'c--Vilonen) in the affine Grassmannian of \( G \) index bases for irreducible representations of \( G^\vee \). In this paper we work with \( G = \GL_m\) and identify \( G^\vee = \GL_m\) as well.

In~\cite{mirkovic2007quiver} (see also the recent sequel~\cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov supply a geometric version of symmetric and skew Howe \((\GL_m,\GL_N)\) dualities.  They relate Kazhdan--Lusztig slices in the affine Grassmannian of \(\GL_m\) to slices in \(N\times N\) nilpotent orbits on the one hand, and to Nakajima quiver varieties on the other hand.    The second author~\cite{dthesis} showed  that under the Mirkovi\'c--Vybornov isomorphism, the MV cycles are identified with certain varieties of matrices, called generalized orbital varieties.

Essential to the geometric Satake equivalence, the BD Grassmannian (short for Beilinson--Drinfeld) is used to define a fusion product on the category of spherical perverse sheaves on the affine Grassmannian.  In~\cite{anderson2003polytope}, Anderson used the BD Grassmannian to define a fusion of MV cycles.  He constructed a family where the general fibre is a Cartesian product of two MV cycles, and the special fibre, called the fusion, is a (non-reduced) union of MV cycles.  Anderson conjectured that this fusion product matches the multiplication in the coordinate ring of the unipotent subgroup \( N \subset\GL_m\).  This conjecture was proven in~\cite{baumann2019mirkovic}, where it was also established that the MV cycles give a basis for \( \CC[N]\).

\subsection{Fusion using generalized orbital varieties}\label{ss:fugov}
% 
Computing the fusion product of MV cycles using the BD Grassmannian is quite difficult, both computationally and conceptually.  A few such computations were done by Anderson and Kogan in~\cite{anderson2006algebra} in a somewhat ad-hoc fashion.  Another perspective on these computations (using convolution Grassmannians) was recently studied by Baumann, Gaussent and Littelmann~\cite{baumann2020bases}.

@@

\begin{theorem}\label{thm:main}
    Let \( Z', Z'' \) be MV cycles. The structure constants for the of the corresponding MV basis vectors in \(\CC[N]\) are equal to the intersection multiplicities of generalized orbital varieties in the fusion of the generalized orbital varieties corresponding to \( Z', Z''\).
\end{theorem}

@@

\subsection{Cluster algebras and MV cycles}\label{ss:clmv}
% 
One motivation for computing products of MV basis elements is to relate the cluster algebra structure on \(\CC[N]\) to the MV basis. Cluster algebras were introduced by Fomin and Zelevinsky in~\cite{fomin2002cluster1} and it was shown in~\cite{geiss2007initial}, using a construction of~\cite{berenstein2005cluster3}, that \(\CC[N]\) has a natural cluster algebra structure.

This structure starts with an initial seed \(\Sigma = (\{x_1,\dots,x_r\},B)\), where \(x_i \in \CC[N]\) are a cluster of variables, \(B\) is a certain matrix, and \( r = m(m-1)/2 \).
We obtain new variables through a process called mutation, where for a mutable variable \(x_i\), its mutation is the unique element \(x_i^* \in \CC[N]\) such that 
\[
    x_i x_i^* = x_+ + x_- \,.
\]
This equation is called an exchange relation, and the terms \(x_+\) and \(x_-\) are certain monomials in the \(x_j\), \(j\neq i\),  determined by \(B\). 
It gives us the new cluster \(\{x_1, \dots, x_i^*, \dots, x_r\}\).

The initial cluster \(\{x_1,\dots,x_r\}\) as well as all possible variables that are obtained through successive mutations are known as cluster variables. 
The cluster monomials are products of cluster variables that are supported on a single cluster.

There are two well-known bases of \(\CC[N]\) whose relationship with the cluster structure has been heavily studied: the dual canonical basis of~\cite{lusztig1990canonicalbases} and the dual semicanonical basis of~\cite{lusztig2000semicanonical}.
In~\cite{kang2018monoidal}, Kang, Kashiwara, Kim and Oh showed that the cluster monomials are contained in the dual canonical basis, while, in~\cite{geiss2006rigid}, Gei\ss, Leclerc and Schr\"oer proved that they are all contained in the dual semicanonical basis.
% \cite[Theorem~2.8(3)]{geiss2006rigid}.
%

The MV basis, the dual canonical basis, and the dual semicanonical basis are all examples of biperfect bases of \(\CC[N]\)~\cite[Section 2]{baumann2019mirkovic}. In particular, they share a crystal structure.
In~\cite[Appendix]{baumann2019mirkovic}, we showed that the MV basis differs from the dual semicanonical basis at the same point in the underlying crystal as where the dual semicanonical basis differs from the dual canonical basis.  
% 
The fact that this point at which each of these bases differ is not a cluster monomial in any of these bases, gives evidence for the following conjecture of Anderson and Kogan~\cite[Conjecture 5.1]{anderson2006algebra}.
\begin{conjecture}\label{conj:cluster in MV}
    The cluster monomials are contained in the MV basis. 
\end{conjecture}

The conjecture is true in type \(A_n\) for \(n\leq 3\) since in these cases \(\CC[N]\) has a unique biperfect basis \cite[Section 2.3]{baumann2019mirkovic}.
% 
More generally, cluster variables called flag minors, coming from initial seeds constructed in \cite[Definitions~2.2--2.3]{berenstein2005cluster3}, belong to each biperfect basis of \(\CC[N]\) \cite[Remark~2.10]{baumann2019mirkovic}. 
And recently in \cite[Proposition 7.2]{baumann2020bases} cluster monomials supported on clusters of flag minors satisfying a certain combinatorial condition where shown to be in the MV basis.

Our hope is that each exchange relation can be realized using the fusion of MV cycles. This would imply that all cluster variables are in the MV basis. In \Cref{ss:GL4 examples} we verify that this is indeed the case in type \(A_3\) by directly computing each fusion product corresponding to an exchange relation.

\section{Supporting Cast}\label{s:players}
% 
\subsection{Rings and discs}\label{ss:rings}
% 
Set \(\cO = \cO_0 = \CC\xt\) and \(\cK = \cK_0 = \CC\xT\).
We will also consider \(\cO_s = \CC\xt[t-s]\) and its fraction field \(\cK_s=\CC\xT[t-s]\), for any \( s \in \CC\setminus\{0\} \), as well as \( \Oinf = \CC\xt[t^{-1}] \) and \(\Kinf = \CC\xT[t^{-1}]\). For any point \( s \in \PP = \PP^1\), \( \cO_s\) is the completion of the local ring \( \mathcal O_{\PP, s} \) and thus the formal spectrum of \( \cO_s\) is the formal neighbourhood \( D_s\) of \( s \) (also called the formal disc centered at \( s\)).  Similarly, the formal spectrum of the field \(\cK_s\) is the deleted formal neighbourhood (or punctured disc), denoted \( D_s^\times\).

Note that for \( s \in \AA = \AA^1\), we have an obvious isomorphism  \(\cO_s \cong \cO\) and \(\cK_s\cong\cK\) taking \( t-s \) to \( t\). 
%

\subsection{Groups}\label{ss:groups}
% 
Let \(H \) be an algebraic group over \( \CC \).  We will be interested in \( H(R)\) where \( R \) is a \(\CC\)-algebra, for example \(R = \CC[t], \cO_s\), etc.  
% 
Note that evaluation at \( t = s\) provides a group homomorphism \( H(\cO_s) \rightarrow H\).  We denote the kernel of this map by \( H_1(\cO_s)\), often called the  first congruence subgroup.  We will be particularly interested in this construction in the case \( s = \infty\), which gives us the group \( H_1(\Oinf)\).

Throughout the paper, we fix \( m \in \NN\).  We let \(G = \GL_m\) and we let \(T\subset G\) be the maximal torus of diagonal matrices. Let \( N, N_- \subset G\) denote the subgroups of upper and lower triangular matrices with 1s on the diagonal. 
We identify \(\ZZ^m\) with the coweight lattice of \(G\) and say that a coweight \(\nu = (\nu_1,\dots,\nu_m)\) is \new{dominant} if \(\nu_1\ge\cdots\ge\nu_m\) and \new{effective} if \(\nu_j\ge 0\) for all \(j\).
% 
If \(\nu\) is both effective and dominant, then it is a partition of size \(|\nu| = \nu_1 + \cdots + \nu_m\in\NN\).  Let \( Q_+ \subset \ZZ^m\) denote the positive root cone, explicitly we have
\[
    Q_+ = \{ (\nu_1, \dots, \nu_m) : \nu_1 + \cdots + \nu_j \ge 0 \text{ for }j = 1, \dots, m-1 \text{ and } \nu_1 + \dots + \nu_m = 0 \} \,. 
\]
We define a partial order on \( \ZZ^m\) by \( \lambda \ge \mu \) if and only if \( \lambda - \mu \in Q_+ \).

Given \(s\in\CC\) we define \((t-s)^\nu\) to be the diagonal matrix 
\[
\begin{bmatrix}
    (t-s)^{\nu_1} \\
    & (t-s)^{\nu_2} \\ 
    & & \ddots \\
    & & & (t-s)^{\nu_m}
\end{bmatrix} 
\]
which we can view in \(G(K)\) for any ring \( K \) containing \((t-s)^{-1}\) and \(\CC[t]\). For example, \((t-s)^\nu\in G(K)\) for \( K = \Ks\) or \(\CC(t)\).

We will also be interested in the affine space \( M_m\) of \(m\times m\) matrices.  Note that for any \( \CC\)-algebra \( R \), \(G(R)\) consists of those matrices \( g \in M_m(R) \) whose determinant is invertible in \( R\). Thus for example \( (t-s)^\mu\in M_m(\CC[t])\) for all effective \( \mu \in \ZZ^m\) but \((t-s)^\mu\in G(\CC[t])\) if and only if \( \mu = 0 \).
%

\subsection{Lattices}\label{ss:lat}
% 
We will use the lattice model for the affine Grassmannian, so it is useful to recall the following definition. Let \( R \subset K\) be two \(\CC\)-algebras (usually, but not always, \(K\) will be a field). 
Consider \( K^m \) as a \(K\)-module. By restriction \( K^m\) can be viewed as an \(R\)-module.  An \new{\(R\)-lattice} in \(K^m\) is an \(R\)-submodule \( L \subset K^m\) which is a free \(R\)-module of rank \( m \) and satisfies \( L \otimes_R K = K^m \). Equivalently, \( L = \Sp_R(v_1, \dots, v_m)\) where \(v_1, \dots, v_m\) are free generators of \(K^m\). 
%

\( R^m \subset K^m \) is called the standard lattice. The group \(\GL_m(K) \) acts transitively on the set of \(R\)-lattices in \(K^m\), thus giving a bijection between this set and \(\GL_m(K)/\GL_m(R)\), since \( \GL_m(R) \) is the stabilizer of the standard lattice.

We will be particularly interested in \(\CC[t]\)-lattices in \( \CC(t)^m\). Given such a lattice \( L \) and a point \( a \in \CC\), the \new{specialization} of \( L \) at \( a \) is the lattice in \(\cK_a^m\) defined as \( L(a) := L \otimes_{\CC[t]} \cO_a \). 
If \(L(a) = \cO_a^m\) then \(L\) is said to be \new{trivial at \(a\)}. 
For example, the lattice \((t-s)^{-1}\CC[t] \subset \CC(t)\) is trivial at any \(a\ne s\), since \( t-s \) is invertible in \( \cO_a\).

\section{Affine Grassmannians}\label{s:affgrs}
% 
We now define various versions of the affine Grassmannian which will play important roles in this paper. 
% 
Each definition is made group-theoretically and then restated as a moduli space of vector bundles and as a moduli space of lattices.
% 
We also sketch how to pass between descriptions.

In these definitions, \( \Vtriv\) denotes the trivial rank \( m \) vector bundle.

\begin{definition}\label{def:gr}
     The \new{ordinary affine Grassmannian} \(\Gr = G(\cK)/G(\cO)\).
\end{definition}    
It is the moduli space of vector bundles with trivializations,
\[
\Gr = 
    \left \{ 
        (V, \varphi) : \text{\(V\) is a rank \(m\) vector bundle on \( D_0 \), \(\varphi : V \xrightarrow{\sim} \Vtriv \) on \( D_0^\times \)} 
    \right \} \,. 
\]
It is also a moduli space of lattices, 
\[ 
\Gr = 
    \left\{ L \subset \cK^m : \text{ \(L\) is a \(\cO\)-lattice} \right\}\,.
\]
We obtain a lattice from a pair \( (V,\varphi) \) by setting \( L = \Gamma(D_0, V)\) which is embedded into \( \cK^m = \Gamma(D_0^\times, \Vtriv)\) using \( \varphi\).  
On the other hand, to get a lattice from the group-theoretic description \( G(\cK) / G(\cO) \) we set \( L = g \cO^m\) for \( g \in G(\cK)\).
% 
\begin{definition}\label{def:grs}
    For any \( s \in \CC \), the \new{ordinary affine Grassmannian} \(\Gr_s = G(\cK_s)/G(\cO_s)\) \new{at} \( s \). 
\end{definition}  
% 
As above, we have modular and lattice descriptions:
% 
\begin{gather*}
\Gr_s = 
    \left\{ (
        V, \varphi) : \text{\(V\) is a rank \(m\) vector bundle on \( D_s \), \(\varphi : V \xrightarrow{\sim} \Vtriv \) on \( D^\times_s \)} 
    \right\}\,, \\
\Gr_s = 
    \left\{ 
        L \subset \cK_s^m : \text{ \(L\) is a \(\cO_s\)-lattice} 
    \right\} \,. 
\end{gather*}
% 
\begin{definition}\label{def:grth}
    The \new{thick affine Grassmannian} \(\Grth = G(\Kinf)/G(\CC[t])\).
\end{definition}
Again, we have the following modular and lattice descriptions:
\begin{gather*}
\Grth = 
    \left\{ 
        (V, \varphi) : \text{\(V\) is a rank \(m\) vector bundle on \( \PP \), \(\varphi : V \xrightarrow{\sim} \Vtriv \) on \( D_\infty \) } 
    \right\}\,, \\
\Grth = 
    \left\{ 
        L : L \subset  \Kinf^m \text{ is a \(\CC[t]\)-lattice} 
    \right\}\,.
\end{gather*}
% 
\begin{definition}\label{def:bdgr}
    The two point \new{Beilinson--Drinfeld Grassmannian} \[\pi : \Grbd\to \AA\] with one point fixed at 0, and the second point \(s\in\AA\) varying.
\end{definition}
% 
It is described in modular terms by 
\[
\Grbd = 
    \left\{ 
        (V,\varphi,s) : V\text{ is a rank \(m\) vector bundle on }\PP, \varphi : V \xrightarrow{\sim} \Vtriv \text{ on } \PP \setminus \{0, s\}  
    \right\}\,. 
\]
% 
The fibre of \(\Grbd \to \AA\) over \( s \in \AA \) will be denoted \( \Gr_{0,s} \) and is given by
% 
\[  
    G(\CC[t, t^{-1}, {(t-s)}^{-1}]/G(\CC[t])\,. % := \pi^{-1}(s) 
\]
We also have the lattice descriptions:
% 
\begin{gather*}
\Grbd = 
    \left \{ 
        (L,s) : L \subset  \CC(t)^m\text{ is a \(\CC[t]\)-lattice trivial at any }a \ne 0, s 
    \right \} \,, \\
% , $L$ is 
\Gr_{0,s} = 
    \left \{ 
        L : L \subset  {\CC[t,t^{-1},{(t-s)}^{-1}]}^m \text{ is a \(\CC[t]\)-lattice} 
    \right \} \,.
\end{gather*}
% 
\begin{definition}\label{def:grplus}
The \new{positive part} of \(\Gr \), resp.\ \(\Grth\), is defined by 
\begin{gather*}
    \Gr^+ = \left(M_m(\cO) \cap G(\cK)\right) / G(\cO)\,,\,\, 
    \text{resp.\ } {\Grth}^+ = \left(M_m(\CC[t]) \cap G(\Kinf)\right) / G(\CC[t])\,. 
\end{gather*}
\end{definition}
% 
In modular terms, \(\Gr^+\) (resp.\ \({\Grth}^+\)) is the set of those \( (V, \varphi)\) where \( \varphi : V \rightarrow \Vtriv \) extends to an inclusion of coherent sheaves over \( D_0 \) (resp.\ over \( \PP\)).

In lattice terms, \( \Gr^+\) (resp.\ \({\Grth}^+\)) contains those lattices \(L\) which are contained in \(\cO^m\) (resp.\ \(\CC[t]^m\)).
%
